Generalized Plackett-Luce Likelihoods

### 2023-05-10 inst/jss_4382_rebuttal.tex

S4 P. 7: "Wolf" should be "Wolfe".
\textcolor{blue}{Corrected.}\\ \\

S5 P. 15, maxp(const2020, n=1): What does n do?
\textcolor{blue}{Argument $n$ discussed above under Timing}\\ \\

S6 P. 16, "we see strong evidence for a real decrease in the strength": I'd tone
that down. The model is almost certainly wrong, so it does not measure
the "real decrease." Also, power is unknown, so it is difficult to say what
"evidence" there is against H0. (See, e.g., Gelman, Regression an other
stories, Chapter 4: Time to unlearn what you thought you knew about
statistics. https://statmodeling.stat.columbia.edu/2022/01/27/regression-
and-other-stories-free-pdf/)\\

\textcolor{blue}{I have not seen ROS before and have been studying
  chapter 4 as suggested.  I make some discursive responses
  here.\\ \\ I was fascinated by ``Researcher degrees of freedom''; I
  am familiar with the concept but have not seen this particular, very
  apposite, phrase used to describe it.  I would observe that my work
  is not in the same class as discussed in ``solving the Bible code
  puzzle'' [in which an unacceptably high amount of researcher DoF was
    documented].  My own work is, for example, written to
  reproducibility standards and all data and indeed code held on
  public repositories [github and CRAN], open to scrutiny.  There is
  very little ``wiggle room'' in my approach: the data is unambiguous,
  and the model (Plackett-Luce) is clear and objective.\\ \\ I would
  observe that Gelman uses the frequentist paradigm uncritically, a
  position adopted by a decreasing minority of researchers.  For
  example, the notion of ``bias'' is discussed; but many workers
  (myself included) regard the bias of an estimator as unimportant: a
  biased estimator with a small mean square error is not without
  value.\\ \\I am not sure what relevance ``power'' has here: this is
  the probability of {\em correctly} failing to reject the null when
  it is false (and here, I did not fail to reject my null: perhaps we
  should include an ensemble of Earths and visit a random one on which
  ``I'' failed to reject the null).\\ \\ The comment ``the model is
  almost certainly wrong'' is, I would argue, misleading.  All models
  are wrong.  But some are useful.  My view is that the truth of a
  statistical model [for some theological definition of truth] is a
  poor criterion: no random variable is truly Gaussian, for example;
  but the Student $t$-test is not useless as a result.  The issue is
  whether non-normality materially affects the properties of the test.
  Often this is not the case and the Gaussian approximation is
  adequate: the $t$-test is sufficiently robust to non-normality to be
  useful in a wide range of situations.  Is the $t$-test True?  No.
  Is it Useful?  Yes.  Consistent?  Yes.  Reliable?  Yes.  Easy to
  use?  Yes.  Familiar?  Yes.  Well-understood?  Yes.  Powerful?  Yes.
  The Student $t$-test has value, and so does Plackett-Luce.  My
  position is that the framework set out in the manuscript is
  plausible and intuitive.  I make no claim for absolute
  correctness.\\ \\ But in answer to your question, the small
  $p$-value and large likelihood ratio statistic certainly {\em is}
  evidence for a real decrease.  The hypothesis of no real strength
  change has low likelihood compared with the alternative of a drop in
  Bradley-Terry strength.  Either Plackett-Luce is seriously wrong, or
  PL is correct but some low-probability event has occurred which is
  misleading us.  Alternatively, as I suggest, Mercedes really is
  suffering from some genuine race-affecting issue [porpoising? Lewis
    Hamilton getting older?  unaccustomed F1 budget caps?].  The {\em
    strength} of this evidence is quantified by the small $p$-value
  (here less than 0.002) or alternatively a likelihood ratio of
  $e^{4.722}\simeq 112.4$, surely worthy of the epithet
  ``strong''.}\\ \\

S7 The references are lacking their DOIs.\\ \textcolor{blue}{Done
  (except for two books---for which I have followed APA6---and R
  itself, for which I used the result of {\tt citation()})} \\ \\

2 Review of software\\

The package includes many functions with a similar name. E.g., {\tt
  ordervec2supp()}, {\tt ordervec2supp3()}, {\tt ordervec2supp3a()},
{\tt ordertable2supp()} , {\tt ordertable2supp3()}. Maybe there could
be some overview of which to use for what
purpose.\\ \textcolor{blue}{An overview is given in {\tt
    ordervec2supp()} and {\tt ordervec2supp3()}, which are
  cross-referenced to each other.  Order tables are documented at {\tt
    ordertable.Rd}; order vectors at {\tt ordervec2supp.Rd}.  The
  functions mentioned by the referee all follow a consistent
  nomenclature: as it says in {\tt hyper3.Rd}, ``Functionality for
  {\tt hyper3} objects is generally indicated by adding a 3 to
  function names, eg {\tt gradient()} goes to {\tt gradient3()}''.}

@@

Section editor comments:

The author proposes an extension of the alreadty existing hyper2 package.
This new S3 class allows for some extended likelihood specifications.

\textcolor{blue}{I would agree with this.}\\ \\

Even though this is just a code snippet and the hyper2 package was
already published elsewhere, the author should provide some
introductory elaborations on the Plackett-Luce model. Along these
lines it would also be helpful to give a quick overview of the hyper2
package, and mention other related packages. As it is now it is
difficult for a reader who's not super familiar with PL-models and
hyper2, respectively, to follow the paper. Anyway, I find the runners
example as motivating example very illustrative, and actually like the
examples throughout the paper.\\ \textcolor{blue}{Thank you!  I too
  have enjoyed the runner paradigm.  But I am aware that the new
  likelihood function is novel and likely to be challenging for those
  not familiar with {\tt hyper2}.  I have added some discussion of
  {\tt hyper2} and its design principles.\\ \\I would not agree with
  the characterization as a code snippet.  The {\tt C++} code base for
  {\tt hyper3} functionality (318 lines) is similar to that for {\tt
    hyper2} (396 lines).  But admittedly {\tt hyper3} functionality
  does leverage many of the S3 methods used in {\tt hyper2}.}\\ \\

Some more aesthetic effort should be made in the code file (e.g.,
providing comments). The code looks like a simple purl()
extraction.\\ \textcolor{blue}{The JSS style file forbids comments
  within the code but encourages authors to comment in the normal
  LaTeX text.  Each code chunk is now either preceded by a description
  of its content, or followed with ``Above, we see\ldots'' or
  equivalent.}

### 2023-05-10 inst/jss_hyper3.Rnw
work on issue #153

However, the majority of software is written in the \proglang{R}
computer language~\citep{rcore2022}, which includes extensive
functionality for working with such likelihood functions:
\cite{turner2020} discuss several implementations from a computational
perspective.  The \pkg{BradleyTerry} package \citep{firth2005}
considers pairwise comparisons using \proglang{glm} but cannot deal
with ties or player-specific predictors; the \pkg{BradleyTerry2}
package \citep{turner2012} implements a flexible user interface and
wider range of models to be fitted to pairwise comparison datasets,
specifically simple random effects.  The \pkg{PlackettLuce} package
\citep{turner2020} generalizes this to likelihood functions of the
form of Equation~\ref{PL_like} and applies the Poisson transformation
of~\cite{baker1994} to express the problem as a log-linear model.  The
\pkg{hyper2} package, in contrast, gives a consistent language
interface to create and manipulate likelihood functions over the
simplex ${S}_n=\left\lbrace\left(p_1,\ldots,p_n\right)\left|p_i\geq
0,\sum p_i=1\right.\right\rbrace$.  A further extension in the package
generalizes this likelihood function to functions of ${\mathbf
p}=(p_1,\ldots,p_n)$ with

>>>>

However, the majority of software is written in the \proglang{R}
computer language~\citep{rcore2023}, which includes extensive
functionality for working with such likelihood functions:
\cite{turner2020} discuss several implementations from a computational
perspective.  The \pkg{BradleyTerry} package \citep{firth2005}
considers pairwise comparisons using \proglang{glm} but cannot deal
with ties or player-specific predictors; the \pkg{BradleyTerry2}
package \citep{turner2012} implements a flexible user interface and
wider range of models to be fitted to pairwise comparison datasets,
specifically simple random effects.  The \pkg{PlackettLuce} package
\citep{turner2020} generalizes this to likelihood functions of the
form of Equation~\ref{PL_like} and applies the Poisson transformation
of~\cite{baker1994} to express the problem as a log-linear model.  The
\pkg{hyper2} package, in contrast, gives a consistent language
interface to create and manipulate likelihood functions over the
simplex ${S}_n=\left\lbrace\left(p_1,\ldots,p_n\right)\left|p_i\geq
0,\sum p_i=1\right.\right\rbrace$.  A further extension in the package
generalizes this likelihood function to functions of ${\mathbf
p}=(p_1,\ldots,p_n)$ with
